---
title: "netflix_04202021"
author: "Namita Trikannad"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
# set default knitr chunks
knitr::opts_chunk$set(
  tidy.opts=list(width.cutoff=60),tidy=TRUE,
  echo = FALSE,  # don't print the code chunk
  results = 'hide',
  warning = FALSE,  # don't print warnings
  message = FALSE,  # don't print messages
  fig.width = 8,  # set default width of figures
  fig.height = 6,  # set default height of figures
  fig.align = "center",  # always align figure in center
  fig.pos = "H",  # always plot figure at the exact location of the code chunk
  cache = F)  # don't cache results  #CHANGE THIS TO "FALSE" AFTER YOU'RE DONE !!

#Remove this 
setwd("GitHub/TidyTuesday")
```


```{r}
library(tidyverse)
library(lubridate)
library(tidytuesdayR)
theme_set(theme_minimal())
library(tidytext)
library(maps)
library(fuzzyjoin)
library(ggthemes)
library(ggpubr)
library(scales)
theme_set(theme_minimal())
library(wordcloud)
library(ggrepel)
library(ggwordcloud)
```

```{r}
tt <- tt_load("2021-04-20")

netflix_titles <- tt$netflix_titles


# which countries have the most content?

netflix_titles %>% 
  count(country, sort = T) 
```

```{r}
head(netflix_titles) 

# Where most content is coming from? There are also many movies that have multiple countries associated with them 
netflix_titles %>% 
  filter(!is.na(country)) %>% 
  count(country, sort = T) %>% 
  top_n(n,n = 10)

# What is the range of the years? 1925 to 2021 
netflix_titles %>% 
  count(release_year, sort = T)

netflix_titles %>% 
   count(type)

netflix_titles %>% 
  ggplot(aes(x = release_year, fill = type)) +
  geom_histogram()
 

# Is the show_id unique to the show/movie? Yes 
netflix_titles %>% 
  group_by(title,duration) %>%
  count(show_id) %>% 
  filter(n > 1)
```

Cleaning steps: 
```{r}
netflix_titles <- netflix_titles %>% 
  separate(duration, c("duration", "duration_unit"), sep = " ") %>% 
  mutate(date_added = lubridate::mdy(date_added),
         added_year = lubridate::year(date_added),
         release_decade = 10 * (release_year %/% 10),
         duration = as.numeric(duration)) 

netflix_titles_unlisted <- netflix_titles %>% 
    separate_rows(listed_in, sep = ", ") 

# Movies and shows listed on Netflix produced entirely or partly in India 
netflix_india <- netflix_titles %>%
  filter(grepl("India", country)) 
```

Questions:
1. When was the date for Netflix India entry? 
2. What are the most common genres in Indian Netflix titles?
3. More TV shows or more movies? 
4. Are there certain directors that are listed more often?
5. How does release date vs date added for indian movie titles
6. Innovative genres between India and US 
7. Timeline of Genres in indian netflix
8. Are indian movies getting shorter on Netflix?
9. What are the longest duration genres in Indian movies?

Indian or India related content on Netflix
```{r}
# What is the oldest Indian movie in Netflix? 
netflix_india %>% 
  filter(release_year ==  min(release_year)) %>% 
  distinct(title)

# When was the first Indian movie added on Netflix?
netflix_india %>% 
  filter(date_added ==  min(date_added)) %>% 
  distinct(title)

# What are the most common genres? 
netflix_india %>% 
  count(listed_in = fct_lump(listed_in, 10), sort = T) %>%
  mutate(listed_in = fct_reorder(listed_in, n)) %>%
  ggplot(aes(n, listed_in)) +
  geom_col()

netflix_india %>% 
  ggplot(aes(added_year, release_year, color = type)) +
  geom_point(alpha = 0.5, position = "jitter")


# Duration of Indian movies over the years?

netflix_india %>% 
  filter(type == "Movie",
         release_decade != 1950) %>% 
  # group_by(release_decade, listed_in) %>% 
  # summarise(median_duration = median(duration, na.rm = T)) 
  ggplot(aes(x = as.factor(release_decade), y = duration)) +
  geom_boxplot() +
  labs(x = "Release Decade", y = "Duration(mins)", 
       title = "Indian movies on Netflix: duration over decades") +
  theme(title = element_text(hjust = 0.5, face = "bold"))


# Most frequently occuring actors by year 
netflix_india %>% 
  separate_rows(cast, sep = ", ") %>% 
  filter(!is.na(cast)) %>% 
  count(type, added_year, cast)  %>%
  group_by(type, added_year) %>%
  slice(which.max(n))

# Most frequently occuring directors by decade
netflix_india %>% 
  separate_rows(director, sep = ", ") %>% 
  filter(!is.na(director)) %>% 
  count(type, release_decade, director)  %>%
  group_by(type, release_decade) %>%
  slice(which.max(n))

# Difference between time it takes for indian movies to be listed 
netflix_india %>% 
  separate_rows(director, sep = ", ") %>%
  filter(director %in% c("Lekh Tandon", "Hrishikesh Mukherjee", "Umesh Mehra", "Ashutosh Gowariker", "Dibakar Banerjee", "Anurag Kashyap")) %>%
  mutate(diff_date = added_year - release_year) %>% 
  ggplot(aes(y = fct_reorder(title, diff_date))) +
  geom_segment(aes(x = added_year, xend = release_year, yend = title)) +
  labs( x = " ", y = " ", 
        title = "Time Between Release Date & Date Added To Netflix",
        subtitle = "Movies selected are directed by the most frequently listed directors of the decade")
```

Main plot: 
```{r}
# Most commonly used words to describe Indian content listed on Netflix over the years they were added/release decade 

ind_words <- netflix_india %>% 
  unnest_tokens(word, description, drop = F) %>% 
  mutate(word = ifelse(grepl("indian", word), "india", 
                                    ifelse(grepl("brother's", word), "brother", word))) %>% 
    filter(word != "sets") %>%
  count(added_year, word, sort = T) %>%
  anti_join(stop_words, by = "word") %>% 
  filter(word != "india") #removing some words manually that are particularly descriptors

ind_words %>%  
 filter(n > 1) %>%
  #filter(release_decade != 1950) %>%
  group_by(added_year) %>% 
  top_n(n, n = 10) %>% 
  ggplot (aes(label = word, size = n, color = n)) +
geom_text_wordcloud() +
  facet_grid(~added_year) +
  scale_color_gradient(low = "red", high = "red") +
  scale_size(range = c(4, 10), guide = FALSE) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
   labs(x = " ", y = " ", title = "What are the most common words used to describe Indian content on Netflix?", 
        subtitle = "Year content was added to Netflix | Top most frequent words in descriptions \n",
        caption = "Visualization: @namitatrix | Source: Tidy Tuesday") +
  facet_grid(~added_year, scales = "free") +
  theme(plot.background = element_rect(fill = "#221F1F"),
        panel.background = element_rect(fill = "#221F1F"),
        plot.title = element_text(hjust = 0.5, size = 20, color = "#F5F5F1"),
        plot.subtitle = element_text(hjust = 0.5, size = 12, color = "#F5F5F1"),
        plot.caption = element_text(size = 10,  color = "#F5F5F1"),
        strip.text.x = element_text(
        size = 12, color = "#F5F5F1", face = "bold"
        ))


ggsave("plots/netflix_042221.png")

```


